# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r}
library (ggplot2)
activity <- read.csv ("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
sum_day <- aggregate (activity$steps, list (activity$date), sum)
```
Mean total number of steps taken per day
```{r}
mean (sum_day[, 2], na.rm = TRUE)
```
Median total number of steps taken per day
```{r}
median (sum_day[, 2], na.rm = TRUE)
```
### Histogram of the total number of steps taken each day
```{r}
hist(sum_day[, 2], xlab="Steps per day", main = paste("Histogram of steps taken each day"))
```
## What is the average daily activity pattern?
```{r}
int_day <- aggregate (activity$steps, list (activity$interval), mean, na.rm = TRUE)
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
which.max (int_day[,2])
```
Plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
g <- ggplot (int_day, aes (int_day[, 1], int_day[, 2]))
g + geom_line () + ggtitle ("Average number of steps by 5-minute interval") + ylab("Number of steps") + xlab ("Minutes")
```
## Imputing missing values

Total number of missing values in the dataset
```{r}
sum (is.na(activity[, 1]))
```
Filling miss vulues by mean for that 5-minute interval from other days
```{r}
activity2 <- activity
i <- 1
while (i <= nrow (activity2) )
{
    if (is.na (activity2 [i, 1]))
    {
        
        activity2 [i, 1] <- int_day [int_day[, 1] == activity2 [i, 3], 2]
    }
i <- i+1
}

```
The mean and median total number of steps taken per day.
```{r}
sum_day2 <- aggregate (activity2$steps, list (activity2$date), sum)
#Mean
mean (sum_day2[, 2], na.rm = TRUE)
#Median
median (sum_day2[, 2], na.rm = TRUE)
```
Q: Do these values differ from the estimates from the first part of the assignment?<br>
A: No<br>
Q: What is the impact of imputing missing data on the estimates of the total daily number of steps?<br>
A: The total daily number of steps will increase<br>

### Histogram of the total number of steps taken each day
```{r}

hist(sum_day2[, 2], xlab="Steps per day", main = paste("Histogram of Steps per day after filling miss vulues"))
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity3 <- activity2
j <-1
while (j <= nrow (activity3))
        {
        ## Add day
        activity3[j, 4] <- format (as.Date(activity3[j,2]), "%w")
        j <- j+1
        }
j <-1
while (j <= nrow (activity3))
        {
        #Add weekday or weekend
        week <- function (x)
                {
                if (x == 6 | x == 0) 
                        {
                        return ("weekend")
                        }
                else
                        {
                        return ("weekday")
                        }
                }
        activity3[j, 5] <- week (activity3[j, 4])
        j <- j+1
        }
#Create the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
int_day_we <- aggregate (activity3$steps, list (activity3$interval, activity3$V5), mean, na.rm = TRUE)
```
### Plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
g <- ggplot (int_day_we, aes (int_day_we[, 1], int_day_we[, 3]))
g + geom_line () +facet_grid (Group.2 ~.) + ggtitle ("Average number of steps across all weekday and weekend days") + ylab("Number of steps") + xlab ("Minutes")
```